---
title: "Practise Assignment"
author: "Vishaylin J. Mahadeo"
date: "August 30, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE ,
                      warning = FALSE , 
                      cache = TRUE ,
                      fig.align = 'center' ,
                      fig.path = 'figure/', 
                      fig.height = 8, 
                      fig.width = 8, 
                      fig.ext = 'png')
library(dplyr)
library(readr)
library(tidyr)
```

#Part 1
## Hello Octocat
I love Octocat. She's the coolest cat in town.
! [](https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/images/octocat.png)

#Part 2
```{r quick_look}
#load data set
data("anscombe")

#column titles
names(anscombe)

#show dimensions of data set rows vs. columns
dim(anscombe)

#show first 6 rows
head(anscombe, 6)

#show last 6 rows
tail(anscombe, 6)

#summary statistics for each column
summary(anscombe)
```

#Part 3
```{r xy_plot}
#load data set
data("anscombe")

#scatter plot (with axis titles, red, and squares)
graph <- plot(x = anscombe$x1, y = anscombe$y1,
     main = 'Figure 1: Scatter plot of anscombe showing x1 vs. y1',
     xlab = 'Anscombe (x1)',
     ylab = 'Anscombe (y1)',
     pch = 15 ,
     col = 2 ,
     lty = 2)

#regression line
regline1 <- lm(y1 ~ x1, data = anscombe)
abline(regline1)
```

#Part 4
```{r data_import}
#load data set
df_analgesic <- read_csv('analgesic.csv')
```

```{r data_summary}
#dimensions of data set
dim(df_analgesic)

#names of column titles
names(df_analgesic)

#first 6 rows of data set
head(df_analgesic)

#last 6 rows of data set
tail(df_analgesic)

#summary statistics for each column
summary(df_analgesic)
```

```{r data_tidy}
#gather data from wide to long format
gathered <- gather(data = df_analgesic ,
                   key = measurements ,
                   value = data, Measurement_1:Measurement_3)

#group by individuals
by_id <- group_by(gathered, ID)
by_id

#summarise and calculate mean
report <- summarize(by_id, mean(data))
report
```

#Part 5
## Chicken Weights
```{r chicken_load}
#load data set(chicken weights)
df_chicken <- read_csv('chick-weights.csv')
boxplot(weight ~ feed, data = df_chicken,
        main = 'Boxplot representing different chicken feeds against chicken weights',
        xlab = 'Chicken feed', ylab = 'Chicken weight')
```


**Null hypothesis:** Type of chicken feed has no effect on chicken weights.

**Alternative hypothesis:** Type of chicken feed does have an effect on chicken weights.

**Assumptions:**

1. a = 0.05

2. Unpaired data.

3. Normally distributed, hence Parametric data set.

4. Therefore One-way ANOVA required to test hypothesis.

5. Reject null hypothesis if p < a

```{r chicken_test}
#One way ANOVA test
chicken_test <- aov(weight ~ feed, data = df_chicken)
summary(chicken_test)

#post-hoc test (Pairwise)
chicken_posthoc <- pairwise.t.test(df_chicken$weight, df_chicken$feed , 
                                   p.adjust.method = 'holm' ,
                                   paired = FALSE)
chicken_posthoc

#make a vector of p-values from each of planned comparisons
p <- c('[test 1]' = 0.001, '[Test 2]' = 0.211, '[Test 3]' = 0.013)
chicken_adjust <- p.adjust(p, method = 'holm')
chicken_adjust
```
F (15.37) = 5.94e^-10; p = 0. Therefore reject null hypothesis, and accept alternative.

**Conclusion:**Type of chicken feed does affect chicken weights.

## The Hot Zone
```{r hot_load}
#load data set (hot_zone)
df_hot <- read_csv('gastroenteritis.csv')
head(df_hot)

#Tabulate data set
hot_table <- table(df_hot$Consumption, df_hot$Outcome)
print(hot_table)

#Cross-tabulate data set
hot_cross <- xtabs(~Consumption + Outcome, data = df_hot)
print(hot_cross)
```

**Null Hypothesis:** Drinking contaminated water does not cause gastroenteritis.

**Alternative Hypothesis:** Drinking contaminated water does cause gasteroenteritis.

**Assumptions:**

1. a = 0.05

2. Unpaired data set.

3. Non-parametric.

4. Therefore Chi squared test is required to test hypothesis.

5. Reject null nypothesis if p < a.

```{r hot_test}
#Pearson Chi squared test
hot_test <- chisq.test(hot_table, correct = FALSE)
hot_test
```
X-squared (2, n=1094)= 74.925; p < 2.2e-16. Therefore reject null hypothesis and accept alternative.

**Conclusion:** Drinking contaminated water does cause gastroenteritis.

## Nausea
```{r nausea_load}
#load data set
df_nausea <- read_csv('nausea.csv')
print(df_nausea)

#exclude outlier
nausea_new <- df_nausea[-8, -1]
print(nausea_new)

#plot data set
boxplot(nausea_new, 
        main = 'Boxplot representing data before and after treatment of 5HT3-receptor blocker',
        xlab = 'Treatment', ylab = 'Intensity rating')
```

**Null Hypothesis:** The 5HT3-receptor blocker does not affect intensity of nausea.

**Alternative Hypothesis:** The 5HT3-receptor blocker does affect intensity of nausea.

**Assumptions:**

1. a = 0.05

2. Paired, ordinal data set

3. Parametric.

4. Therfore a Wilcox test is required to test hypothesis.

5. Reject null hypothesis if p < a.

```{r nausea_test}
#Students t-test
nausea_test <- wilcox.test(nausea_new$Nausea_before, nausea_new$Nausea_after,
                           paired = TRUE)
nausea_test
```
p = 0.04983, therefore reject null hypothesis and accept alternative.

**Conclusion:** The 5HT3-receptor blocker does affect intensity of nausea.

# Part 6
## Housing Prices

```{r house_load}
#load data set
df_house <- read_csv('housing-prices.csv')
house_plot <- plot(df_house$interest_rate, df_house$median_house_price_USD,
                   main = 'Scatter plot showing interest rates vs. house prices',
                   xlab = 'Interest rate (%)', ylab = 'House price (USD)',
                   pch = 15 , col = 2)

#Regression line
regline2 <- lm(median_house_price_USD ~ interest_rate, data = df_house)
abline(regline2)
summary(regline2)

#Check homoskedasticity
homosk <- plot(x = regline2$fitted, y = regline2$residuals,
               main = 'Homoskedasticity check',
               xlab = 'Fitted', ylab = 'Residuals',
               pch = 15 , col = 2)
abline (h = 0)

#Check Gaussian distribution
qqnorm(regline2$residuals)
qqline(regline2$residuals)
```

**Null hypothesis:** Interest rate (%) does not affect house prices (USD).

**Alternative hypothesis:** Interest rate (%) does affect house prices (USD).

**Assumptions:**

1. a = 0.05

2. Unpaired, continous data set

3. Rho = 0.2848, therefore weak relationship between interest rate and house price. 

4. Residuals are homoskedasticity because they show zero spread around abline at 0.

5. Data fitted normaly around Q-Q plot, hence Gaussian residual distribution (i.e. normally distributed)

6. Reject null hypothesis if p < a.

7. Therfore a Pearson's correlation is required to analyze data.

```{r house_test}
#Pearson correlation
house_test <- with(df_house,
                   cor.test(x = interest_rate, y = median_house_price_USD,
                            method = 'pearson'))

#print out result
house_test
```
t(14) = -2.6409; r = -0.5766; p = 0.01937. Therefore reject null hypothesis and accept alternative. 

**Conclusion:** Interest rates (%) have a mild negative relationship with house prices (USD).